Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-03-15 paper/tex/discussion.tex
checking references and moving content
moving content

%The goal was not to determine a ``best'' photo-$z$ code.

@@

In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.

\subsection{Interpretation of metrics}
\label{sec:caution}
%Caution on metrics and photo-$z$ codes}\label{sec:caution}
%(Alex Malz, Sam Schmidt)

We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.

The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as a control case to demonstrate this point via \textit{reductio ad absurdum}.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
%\textit{Explain the connection between \trainz\ and these metrics.}

@@

Trends in the metrics shared among codes indicate the

One obvious trend in several of the codes tested was an overall over or underprediction of the widths of \pzpdf s, as evidenced by the QQ plots and PIT histograms shown in Figure~\ref{fig:pitqq}.
A more careful tuning of bandwidth or smoothing during the validation process appears to be necessary for many of the machine learning based codes in order to improve the accuracy of $p(z)$.
For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.

Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.
Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.
Specific metric choice can affect which codes are among the ``best'' codes.
As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.

@@

\subsubsection{Nontrivial experimental design}
\label{sec:futureexperiments}

@@

The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.
This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
As mentioned in Section~\ref{sec:metrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

@@

\subsubsection{More realistic mock data}
\label{sec:futuredata}

>>>>

@@

In contrast with other \pzpdf\ comparison papers that have aimed to identify the ``best'' code for a given survey, we have focused on the somewhat more philosophical questions of how to assess \pzpdf\ methods and how to interpret differences between codes in terms of \pzpdf\ performance.
In Section~\ref{sec:caution}, we reframe the strong performance of our pathological \pzpdf\ technique, \trainz, as a cautionary tale about the importance of choosing appropriate comparison metrics.
In Section~\ref{sec:newcodes}, we interpret the performance of the genuine \pzpdf\ codes in terms of concrete improvements that should be pursued by the code developers.
In Section~\ref{sec:futureexperiments}, we outline the experiments we intend to build upon this study
In Section~\ref{sec:futuredata}, we discuss the enhancements of the mock data set that will be necessary to enable the future experiments.

\subsection{Interpretation of metrics}
\label{sec:caution}
%(Alex Malz, Sam Schmidt)

We remind the reader that contributed codes were given a goal of obtaining accurate \pzpdf s, not an accurate stacked estimator of the redshift distribution, so we do not expect the same codes to necessarily perform well for both classes of metrics.
Indeed, the codes were optimized for their interpretation of our request for ``accurate \pzpdf s,'' and we expect that the implementations would have been adjusted had we requested optimization of the traditional metrics of Appendices~\ref{sec:moments, sec:pointmetrics}.

Furthermore, our metrics are not necessarily able to assess the fidelity of individual \pzpdf s relative to true posteriors.
Metric-specific performance implies that we may need multiple \pzpdf\ approaches tuned to each metric in order to maximize returns over all science cases in large upcoming surveys.

The \trainz\ estimator, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set as described in Section~\ref{sec:trainz}, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
\textit{We make the alarming observation that \trainz\ outperforms all codes on the PIT-based metrics, and all but one code on the $N(z)$ based statistics.}
%\textit{Explain the connection between \trainz\ and these metrics.}

@@

\subsection{Evolution of \pzpdf\ codes}
\label{sec:newcodes}

Trends in the metrics shared among codes indicate avenues for improvement of each approach.

Section~\ref{sec:pitqq} identifies over- and under-estimation of the widths of \pzpdf s by certain codes.
We suggest that the smoothing parameter present in \bpz, \cmnn, \delight, \eazy, \gpz, \skynet, and \tpz\ be chosen more carefully to improve their performance in the QQ plots and PIT histograms of

@@

\subsection{Extended experimental design}
\label{sec:futureexperiments}

@@

The work presented in this paper is only the first step in characterizing current photo-$z$ codes and moving toward an improved photometric redshift estimator.

As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.
We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.

This initial paper explored code performance in idealized conditions with perfect catalog-based photometry and representative training data.
As mentioned in Section~\ref{sec:metrics} for the stacked $N(z)$ metrics we examined only the entire galaxy population with no selections in either photo-$z$ ``quality" or redshift.
The cosmological analyses for weak lensing and large scale structure based measures plan to break galaxy samples into tomographic redshift bins, using photo-$z$ $p(z)$ to infer the redshift distribution for each bin.
The specific selection used to determine these bins, both algorithmically and the specific bin boundaries, could induce biases due to indirect selections inherent in the photo-$z$ or other bin selection parameters.
The effects of tomographic bin selection will be explored in a dedicated future paper, including propagation of redshift uncertainties in a set of fiducial tomographic redshift bins in order to estimate impact on cosmological parameter estimation.

@@

Sequels to this study will also address some shortcomings of our experimental procedure.
For narrow peaked $p(z)$ the parameterization of the PDF as evaluated on a fixed redshift grid could also have contributed to some overestimates of $p(z)$ width simply due to the finite resolution.
After evaluating results such as those presented in \citet[]{Malz:qp}, in future analyses we plan to switch from a fixed grid to quantile-based storage of $p(z)$ in order to more efficiently and accurately store redshift PDF results.

\subsection{Realistic mock data}
\label{sec:futuredata}

### 2019-03-15 paper/tex/results.tex
moving content

We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal redshift solution, a broad unimodal solution, a bimodal distribution, and a multimodal distribution.

@@

The most striking differences between codes are due to small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
The $\rm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.
However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.
The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
%In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.  We will discuss PDF storage in Section~\ref{sec:discussion}.
%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
% We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}

>>>>

We begin with a demonstrative visual inspection of the \pzpdf s produced by each code for individual galaxies.
Figure~\ref{fig:pz_examples} shows the \pzpdf s for four galaxies chosen as examples of \pzpdf\ archetypes: a narrow unimodal PDF, a broad unimodal PDF, a bimodal PDF, and a multimodal PDF.
We reiterate that under our idealized experimental conditions, differences between codes are the isolated signature of the implicit prior due to the method by which the \pzpdf s were derived.

@@

The most striking differences between codes are due to small-scale features induced by the interaction between the shared piecewise constant parameterization of $200$ bins $0 < z < 2$ of Section~\ref{sec:metrics} and the smoothing conditions or lack thereof in each algorithm.
The $\rm{d}z = 0.01$ redshift resolution is sufficient to capture the broad peaks of faint galaxies' \pzpdf s with large photometric errors but is too broad to resolve the narrow peaks for bright galaxies' \pzpdf s with small photometric errors.
This observation is consistent with the findings of \citet[]{Malz:qp} that the piecewise constant parameterization underperforms in the presence of small-scale structures.

However, the shared small-scale features of \annz, \metaphor, \cmnn, and \skynet\ are a result of various weighted sums of the limited number of training set galaxies with colors similar to those of the test set galaxy in question, with behavior closer to classification than regression in the case of \annz.
The settings used on \gpz\ in this work forced broadening of the single Gaussian to cover the multimodal redshift solutions of the other codes.
%Interestingly, while \textsc{ANNz2} shows an abundance of small scale structure in individual $p(z)$ measurements (see Fig.~\ref{fig:pz_examples}), the summed $\hat{N}(z)$ is rather smooth, where the small scale features average out.  This is not the case for the two other codes that show an abundance of substructure in their individual $p(z)$: both \textsc{CMNN} and \textsc{SkyNet} show small scale features both in $p(z)$ and $\hat{N}(z)$.
%In contrast, FlexZBoost, for example, can return estimates on any grid without compression errors as it’s a basis expansion method where only the expansion coefficients need to be stored.
%Codes with a native output format other than the shared piecewise constant binning scheme (or one that can be losslessly converted to it) may suffer from loss of information when converting to it, which could artificially favor some codes over others in a limited number of cases, for example bright galaxies with very narrow $p(z)$ where the true peak falls between grid points.  We will discuss PDF storage in Section~\ref{sec:discussion}.
%Furthermore, the fidelity of photo-$z$ interim posteriors in this format varies with the quality of the photometry.
%Switching to a quantile based parameterization may be more costly computationally, for example template-based codes would need to test more grid point in order to resolve the quantiles for bright galaxies.  However, the computational time for template based codes scales roughly linearly with the number of grid points, so this may be a reasonable thing to implement.
% We will discuss this further in Section~\ref{sec:discussion}.
% \red{someone review this statement to make sure that I'm saying this correctly!}
